The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization

### 2025-11-27 article_final/main.tex
New style file reached

%\documentclass[wcp,gray]{jmlr} % test grayscale version
 %\documentclass[wcp]{jmlr}% former name JMLR W\&CP
\documentclass[pmlr]{jmlr}% new name PMLR (Proceedings of Machine Learning)

\RequirePackage{graphicx}
 % The following packages will be automatically loaded:
 % amsmath, amssymb, natbib, graphicx, url, algorithm2e
 \usepackage{booktabs}
 %\usepackage{rotating}% for sideways figures and tables
\usepackage{longtable}% for long tables
\usepackage{arydshln}% for dashed lines in tables

@@

% Custom math commands
% Blackboard bold
\newcommand{\R}{\mathbb{R}}

% Bold matrices
\newcommand{\mA}{\bm{A}}
\newcommand{\mB}{\bm{B}}
\newcommand{\mD}{\bm{D}}
\newcommand{\mF}{\bm{F}}
\newcommand{\mG}{\bm{G}}
\newcommand{\mM}{\bm{M}}
\newcommand{\mN}{\bm{N}}
\newcommand{\mS}{\bm{S}}
\newcommand{\mU}{\bm{U}}
\newcommand{\mV}{\bm{V}}
\newcommand{\mX}{\bm{X}}
\newcommand{\mSigma}{\bm{\Sigma}}

% Vectors (note: \vec is already defined by jmlr class)
% We define shortcuts for commonly used vectors
\newcommand{\vu}{\vec{u}}
\newcommand{\vv}{\vec{v}}

% Calligraphic letters
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cS}{\mathcal{S}}

% Norms and operators
\newcommand{\norm}[1]{\lVert #1\rVert}
\DeclarePairedDelimiter{\normf}{\|}{\|_\mathrm{F}}
\DeclarePairedDelimiter{\norms}{\|}{\|_{\mathrm{2}}}
\DeclarePairedDelimiter{\normc}{\|}{\|_{\mathrm{C}}}
\DeclarePairedDelimiter{\normn}{\|}{\|_{*}}
\newcommand{\normkfk}[1]{\|#1\|_{\text{KF-}k}}
\DeclarePairedDelimiter{\normfstar}{\|}{\|_\mathrm{F*}}
\DeclarePairedDelimiter{\normftwo}{\|}{\|_\mathrm{F2}}

@@

% Math operators
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\diag}{diag}

@@

\begin{table}[ht]
	\centering
	\begin{tabular}{|c|c|c|c|c|c|c|c|}
	  \hline
	  Matrix sizes & $k$  & Method           & Time (s) & Matvecs & Iterations & $err_1$         & $err_2$       \\
	  \hline
	  $500 \times 500$      & 5  & Power Iterations           & 0.062   & 2005    & 200        & 9.2$\cdot 10^{-3}$ & 9.1$\cdot10^{-3}$\\
	  $500 \times 500$      & 5  & RSVD             & 0.017   & 1170    & 38         & 9.8$\cdot 10^{-3}$ & 9.6$\cdot10^{-3}$\\
	  $500 \times 500$      & 5  & TRLan            & 0.018   & 131     & 65         & 9.6$\cdot 10^{-5}$ & 9.4$\cdot10^{-5}$\\
	  \hdashline
	  $500 \times 500$      & 50 & Power Iterations           & 0.44    & 43750   & 437      & 9.9$\cdot 10^{-3}$ & 9.0$\cdot10^{-3}$\\
	  $500 \times 500$      & 50 & RSVD             & 0.61    & 6120    & 50      & 9.9$\cdot 10^{-3}$ & 9.1$\cdot10^{-3}$\\
	  $500 \times 500$      & 50 & TRLan            & 0.16    & 462     & 231      & 3.3$\cdot 10^{-7}$ & 3.0$\cdot10^{-7}$\\
	  \hdashline
	  $5000 \times 5000$    & 5  & Power Iterations           & 9.6     & 9065    & 906      & 8.6$\cdot 10^{-3}$ & 8.6$\cdot10^{-3}$\\
	  $5000 \times 5000$    & 5  & RSVD             & 2.1     & 5640    & 187      & 9.7$\cdot 10^{-3}$ & 9.7$\cdot10^{-3}$\\
	  $5000 \times 5000$    & 5  & TRLan            & 0.70    & 205     & 102      & 7.7$\cdot 10^{-3}$ & 7.7$\cdot10^{-3}$\\
	  \hline
	\end{tabular}
  \caption{Comparison of methods for computing rank-$k$ updates on dense random matrices (CPU, double precision). Lower is better in all columns.}
  \label{tbl:matrix_methods}
  \end{table}

  \begin{table}[ht]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
  \hline
  Matrix size    & Time (s) & Matvecs & Iterations & $err_1$ \\
  \hline
  $500 \times 500$     & 0.041    & 27 000  & 27         & 4.8e-3  \\
  $5000 \times 5000$   & 26.4     & 290 000 & 29         & 6.5e-3  \\
  \hline
  \end{tabular}
  \caption{Newton–Schulz iteration on the same matrices (for reference).}
  \label{tbl:newton_schulz_reference}
  \end{table}

@@

>>>>

%\documentclass[wcp,gray]{jmlr} % test grayscale version
 %\documentclass[wcp]{jmlr}% former name JMLR W\&CP
\documentclass[pmlr]{jmlr}% new name PMLR (Proceedings of Machine Learning)
\input{math_commands.tex}
\RequirePackage{graphicx}
 % The following packages will be automatically loaded:
 % amsmath, amssymb, natbib, graphicx, url, algorithm2e
 \usepackage{booktabs}
 %\usepackage{rotating}% for sideways figures and tables
\usepackage{longtable}% for long tables
\usepackage{xcolor}% required by colortbl
\usepackage{colortbl}% for colored table rules (required by arydshln)
\usepackage{arydshln}% for dashed lines in tables

@@

% Norms and operators
\newcommand{\norm}[1]{\lVert #1\rVert}
\DeclarePairedDelimiter{\normf}{\|}{\|_\mathrm{F}}
\DeclarePairedDelimiter{\norms}{\|}{\|_{\mathrm{2}}}
\DeclarePairedDelimiter{\normc}{\|}{\|_{\mathrm{C}}}
\DeclarePairedDelimiter{\normn}{\|}{\|_{*}}
\newcommand{\normkfk}[1]{\|#1\|_{\text{KF-}k}}
\DeclarePairedDelimiter{\normfstar}{\|}{\|_\mathrm{F*}}
\DeclarePairedDelimiter{\normftwo}{\|}{\|_\mathrm{F2}}
\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}

@@

% Math operators
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\diag}{diag}

@@

\begin{table}[ht]
\floatconts
  {tbl:matrix_methods}
  {\caption{Comparison of methods for computing rank-$k$ updates on dense random matrices (CPU, double precision). Lower is better in all columns.}}
  {%
    \begin{tabular}{|c|c|c|c|c|c|c|c|}
      \arrayrulecolor{black}\hline
      Matrix sizes & $k$  & Method           & Time (s) & Matvecs & Iterations & $err_1$         & $err_2$       \\
      \arrayrulecolor{black}\hline
      $500 \times 500$      & 5  & Power Iterations           & 0.062   & 2005    & 200        & 9.2$\cdot 10^{-3}$ & 9.1$\cdot10^{-3}$\\
      $500 \times 500$      & 5  & RSVD             & 0.017   & 1170    & 38         & 9.8$\cdot 10^{-3}$ & 9.6$\cdot10^{-3}$\\
      $500 \times 500$      & 5  & TRLan            & 0.018   & 131     & 65         & 9.6$\cdot 10^{-5}$ & 9.4$\cdot10^{-5}$\\
      \hdashline
      $500 \times 500$      & 50 & Power Iterations           & 0.44    & 43750   & 437      & 9.9$\cdot 10^{-3}$ & 9.0$\cdot10^{-3}$\\
      $500 \times 500$      & 50 & RSVD             & 0.61    & 6120    & 50      & 9.9$\cdot 10^{-3}$ & 9.1$\cdot10^{-3}$\\
      $500 \times 500$      & 50 & TRLan            & 0.16    & 462     & 231      & 3.3$\cdot 10^{-7}$ & 3.0$\cdot10^{-7}$\\
      \hdashline 
      $5000 \times 5000$    & 5  & Power Iterations           & 9.6     & 9065    & 906      & 8.6$\cdot 10^{-3}$ & 8.6$\cdot10^{-3}$\\
      $5000 \times 5000$    & 5  & RSVD             & 2.1     & 5640    & 187      & 9.7$\cdot 10^{-3}$ & 9.7$\cdot10^{-3}$\\
      $5000 \times 5000$    & 5  & TRLan            & 0.70    & 205     & 102      & 7.7$\cdot 10^{-3}$ & 7.7$\cdot10^{-3}$\\
      \arrayrulecolor{black}\hline
    \end{tabular}%
  }
\end{table}

\begin{table}[ht]
\floatconts
  {tbl:newton_schulz_reference}
  {\caption{Newton–Schulz iteration on the same matrices (for reference).}}
  {%
    \begin{tabular}{|c|c|c|c|c|}
      \arrayrulecolor{black}\hline
      Matrix size    & Time (s) & Matvecs & Iterations & $err_1$ \\
      \arrayrulecolor{black}\hline
      $500 \times 500$     & 0.041    & 27 000  & 27         & 4.8e-3  \\
      $5000 \times 5000$   & 26.4     & 290 000 & 29         & 6.5e-3  \\
      \arrayrulecolor{black}\hline
    \end{tabular}%
  }
\end{table}

\section{Experiments}\label{sec:experiments}

\subsection{Randomized Linear Least Squares}
\label{subsec:lls_exps}

First, we compare F-Fanions on the following convex $L$-smooth problem:
\begin{equation}\label{eq:lls}
  F(\mX) = \frac{1}{2} \<(\mX-\mS), \mM (\mX - \mS) \mN> \to \min_{\mX \in \Rmn}\,
\end{equation}
where $\mX \in \Rmn$, $m=500$ and $n=500$ are typical dimensions of a neural network weight matrix, $\mS \in \Rmn$, $\mM \in \mathbb{S}^m_{+}$ and $\mN \in \mathbb{S}^n_{+}$ are positive-semidefinite matrices. The spectra of $\mM$ and $\mN$ are uniformly distributed in the $(0, 1)$ interval. We set $\mS=0$, and $\mX^0 \sim 0.1 \cN(0, \mI)$.

We run different Fanions and their respective F-Fanions with $\alpha=1/2$: Neon (Fanion-1), Fanion-2, Fanion-10, Fanion-100, and Muon (Fanion-500). Also, we test S-Fanions with $\alpha=1/2$ and \texttt{sign\_lr\_coeff=0.01}. We test them against Normalized SGD and SignSGD, which are also F-Fanion and S-Fanion respectively with $\alpha = 0$ and an arbitrary $k$.

Since theoretical bounds \citep{kovalev2025understanding, riabinin2025gluon} rely on a very loose norm bound $\norm{\cdot}\le \rho \normf{\cdot}$, we do not derive learning rate or Nesterov momentum from the smoothness constants that also depend on the norm choice. Rather, we find such \texttt{(lr, momentum)} pair that the loss threshold $0.001$ is reached in a minimal number of iterations. This setting is the most realistic, and at the same time aligns with the corollaries of the convergence theorems where constant learning rate and momentum coefficients are proposed.

We perform a grid search for \texttt{momentum} $\in \{0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95\}$ and for \texttt{lr}. For Muon's, F-Muon's, S-Muon's \texttt{lr} $\in [0.005, 0.0020]$. For SignSGD, \texttt{lr} $\in 0.01 \times [0.005, 0.020]$, and for NSGD \texttt{lr} $\in [0.01, 0.10]$. The tuned parameters and the number of iterations to converge to $0.001$ are in \tableref{tbl:lls_lrs}. \texttt{lr} and \texttt{momentum}, and \texttt{sign\_lr\_coeff} for Fanions, F-Fanions, and S-Fanions are the same as for Muon, F-Muon, and S-Muon respectively due to the impossibility to tune them properly: their losses decrease too slow to reach 0.001.

The results are presented in \figureref{fig:lls} and detailed in \sectionref{sec:lls_plots_section}.

Both F-Muon and S-Muon converge faster and to a lower loss and a lower Frobenius norm of the gradient than NSGD, Muon, or SignSGD.

\begin{table}[htbp]
\floatconts
  {tbl:lls_lrs}
  {\caption{Tuning learning rates and momentum coefficients}}
  {%
    \begin{tabular}{|c|c|c|c|}
      \arrayrulecolor{black}\hline
      Algorithm & \texttt{lr}  & \texttt{momentum}  & Iterations to 0.001 loss \\
      \arrayrulecolor{black}\hline
      Muon & 0.007 & 0.5 & 1060 \\
      \hdashline
      NSGD & 0.08 & 0.95 & 1020 \\
      F-Muon & 0.015 & 0.7 & 910 \\
      \hdashline
      SignSGD & 0.016 $\times$ 0.01 & 0.95 & 2650 \\
      S-Muon & 0.011 & 0.9 & 890 \\
      \arrayrulecolor{black}\hline
    \end{tabular}%
  }
\end{table}

\begin{figure}[htbp]
\floatconts
  {fig:lls}
  {\caption{Linear least squares problem for a 500x500 matrix}}
  {%
    \subfigure[The loss]{\label{fig:lls_loss}%
      \includegraphics[width=0.8\linewidth]{figs/simple_lls/loss_long_run_16_increase.pdf}}%
    \vspace{1em}
    \subfigure[The Frobenius norm of the gradient]{\label{fig:lls_fro_grad_norm}%
      \includegraphics[width=0.8\linewidth]{figs/simple_lls/gradient_frobenius_norm_vs_iteration_long_run.pdf}}
  }
\end{figure}

\subsection{CIFAR-10 airbench}

We compare the algorithms on the CIFAR-10 airbench \citep{cifar2023airbench}. First, to test the impact of $\alpha$, we run F-Muon for different $\alpha$ with the same \texttt{lr=0.24(1 - step/total\_steps), momentum=0.65, nesterov=True} and weight normalization, as have been finetuned by Keller Jordan, 10 repetitions for each $\alpha$. We record the accuracy after 8 epochs of training (\figureref{fig:muon_alphas}).

Then we tune F-Muon with $\alpha = 0.5$. Tuned parameters are \texttt{lr=0.4(1 - step/total\_steps), momentum=0.6, nesterov = True}. Tuned F-Muon reaches $94.02 \pm 0.13\%$ after 8 epochs (averaged by 200 iterations), matching Muon's accuracy and variance. Once again we measure \texttt{val\_accuracy}($\alpha$) curve (\figureref{fig:fmuon_alphas}). We notice that even when $\alpha=0.1$, the accuracy is much higher than in the case of vanilla NSGD.

We tune S-Muon with $\alpha=0.5$ for the best \texttt{(lr, momentum, sign\_lr\_coeff)} and by setting \texttt{lr=0.42(1 - step/total\_steps), momentum=0.63, nesterov = True, sign\_lr\_coeff = 0.003} get $\mathbf{94.03} \pm 0.13\%$, which is even higher than for vanilla Muon.

\begin{figure}[htbp]
\floatconts
  {fig:diff_alphas}
  {\caption{Mean validation accuracies for F-Muon with different $\alpha$}}
  {%
    \subfigure[With parameters tuned for Muon]{\label{fig:muon_alphas}%
      \includegraphics[width=0.48\linewidth]{figs/muon_tuned_diff_alpha.pdf}}%
    \hfill
    \subfigure[With parameters tuned for F-Muon]{\label{fig:fmuon_alphas}%
      \includegraphics[width=0.48\linewidth]{figs/fmuon_tuned_diff_alpha.pdf}}
  }
\end{figure}

The Muon-level performance is surprising for the contorted geometry of F-Muon and S-Muon. F-Muon can be visualized by \figureref{fig:cifar_ball}, where the LMO balls of Muon and F-Muon for real \texttt{lr} are plotted in a 2D space of singular values. S-Muon cannot be visualized because the $l_\infty$ norm is not a function of singular values, while $\norms{\cdot}$, when plotted for the $\Rmn$ space with $m, n > 2$, requires at least a 4D space. However, it is plain that the LMO ball of S-Muon significantly differs from the ball of Muon because the $\texttt{lr}(1-\alpha)\texttt{sign\_lr\_coeff} = 6.3 \times 10^{-4}$, which is comparable to the usual \texttt{lr} of SignSGD for deep learning problems.

The most pathological case is $\alpha > 1$, which corresponds to the LMO with an area that is not a norm ball! Despite this violation, the mixture of algorithms reaches almost the same accuracy as vanilla Muon.

These observations raise the question of how much LMO-based algorithms are sensitive to the constraint area.

\begin{figure}[htbp]
\floatconts
  {fig:cifar_ball}
  {\caption{Visualization of the LMO balls for Muon and F-Muon.}}
  {\includegraphics[width=0.6\linewidth]{figs/norms/fstardual_cifar.pdf}}
\end{figure}

\subsection{NanoGPT speedrun}

We test F-Muon on the NanoGPT speedrun \citep{modded_nanogpt_2024}. For $\alpha = 0.5$, the optimal parameters are \texttt{lr=0.07, momentum=0.95}, while for Muon they were \texttt{lr=0.05, momentum=0.95}. After testing for 1750 steps we get 3.281 cross-entropy loss. For Muon, it falls below the target threshold 3.28 and reaches 3.279. However, this difference is negligible if one looks at \figureref{fig:gpt_loss}. It is even more striking, considering the fact that this F-Muon is an average of Muon and NSGD, and the latter performed quite poorly.

\begin{figure}[htbp]
\floatconts
  {fig:gpt_loss}
  {\caption{The validation loss for NanoGPT}}
  {\includegraphics[width=0.7\linewidth]{figs/nanogpt/val_loss_vs_step.pdf}}
\end{figure}

Again, as on the CIFAR airbench, if we set $\alpha=-0.1$ that corresponds to no ball, we get a 3.2818 loss, which is a small difference from the baseline.

\subsection{NanoGPT Fine-Tuning}
\label{subsec:finetune}

We fine-tuned the NanoGPT framework by Karpathy \citep{Karpathy2022} using the standard \textit{NanoGPT-medium} configuration, which corresponds to GPT-2 Medium. This model consists of 24 transformer layers, a hidden size of 1024, 16 attention heads, and approximately 345 million parameters. The training dataset was the \textit{Tiny Stories} corpus \citep{eldan2023tinystoriessmalllanguagemodels}, selected for its high entropy and structural diversity, which amplify and make more visible the differences in optimization dynamics. All training runs were initialized with the same random seed, weight initialization, and learning rate schedule to ensure that any differences in performance arose solely from the choice of optimizer.

For all one-dimensional layers, AdamW was used with a learning rate of $1\!\times\!10^{-3}$. Since momentum exhibited negligible influence on training dynamics, it was held constant across all experiments. A comparative analysis of optimization performance is provided in \figureref{fig:nanogptftexps}.

\begin{figure}[htbp]
\floatconts
  {fig:nanogptftexps}
  {\caption{Comparison of Muon, F-Muon, and S-Muon across a range of learning rates. Stars denote the best learning rate identified for each optimizer. Validation loss.}}
  {\includegraphics[width=\linewidth]{figs/nanogpt/optimizers_comparison.pdf}}
\end{figure}

\figureref{fig:nanogptftexps2} presents the train and validation loss curves at the optimal learning rate for each optimizer. While F-Muon and S-Muon maintain consistent behavior, vanilla Muon attains the lowest loss overall.

\begin{figure}[htbp]
\floatconts
  {fig:nanogptftexps2}
  {\caption{Train and validation loss at the optimal learning rate for each optimizer. The standard Muon method achieves the lowest loss, while F-Muon and S-Muon remain stable but exhibit slightly inferior performance.}}
  {\includegraphics[width=\linewidth]{figs/nanogpt/Nanogpt finetune results.pdf}}
\end{figure}

\section{Algorithms for Vectors $\rightarrow$ Algorithms for Matrices}

One cannot but notice the similarity between the updates of LMO optimizers in the vector $l_p$ and the matrix Schatten $S_p$ norms, which is illustrated by \tableref{tbl:mat_vs_vec_lmo}. The parallels are not limited to mere similarity of the updates and can be traced even in the empirical observations. SignSGD is very close to Adam, as noted in \citet{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well in training large models \citep{zhao2025deconstructing, liu2025muon}. NSGD is the same for both matrix and vector cases. Greedy Coordinate Descent is not applied to high-dimensional problems, so from this perspective it is not surprising that one-rank Neon underperforms on such problems.

\begin{table}[t]
\floatconts
  {tbl:mat_vs_vec_lmo}
  {\caption{LMO optimizers in Schatten $S_p$ norms and in $l_p$ norms. $g$ is the gradient. When it is a matrix, $g = \mU \mSigma \mV^\top$}}
  {%
    \begingroup
    \def\arraystretch{1.2}
    \resizebox{\linewidth}{!}{%
      \begin{tabular}{|l|c|c|c|}
        \arrayrulecolor{black}\hline
        Algorithm                             & LMO constraint set $\mathcal{D}$ & LMO                                                         & Reference                            \\
        \arrayrulecolor{black}\hline
        \arrayrulecolor{black}\hline
        Normalized SGD                     & $l_2$-ball, $S_2$-ball          & $-\eta \tfrac{g}{\norm{g}_2} = -\eta \tfrac{g}{\normf{g}}$ & \citep{hazan2015beyond}              \\
        Momentum Normalized SGD            & Ball in $l_2$, or Ball in $S_2$ & $-\eta \tfrac{g}{\norm{g}_2} = -\eta\tfrac{g}{\normf{g}}$  & \citep{cutkosky2020momentum}         \\
        \arrayrulecolor{black}\hline
        SignSGD                            & Ball in Max-norm $l_\infty$     & $-\eta \sign(g)$                                            & \citep[Thm. 1]{bernstein2018signsgd} \\
        Signum                             & Ball in Max-norm $l_\infty$     & $-\eta \sign(g)$                                            & \citep[Thm. 3]{bernstein2018signsgd} \\
        \hdashline
        Muon                               & Ball in Spectral $S_\infty$     & $-\eta \mU\mV^\top$                                             & \citep{jordan2024muon}               \\
        \arrayrulecolor{black}\hline
        Gauss-Southwell Coordinate Descent & Ball in $l_1$                   & $-\eta \sum_{i \in \argmax|g_i^t|} \sign(g_i^t)e_i$                       & \citep[p. 19]{shi2016primer}         \\
        \hdashline
        Neon                               & Ball in Nuclear $S_1$           & $-\eta \vu_1 \vv_1^\top$                                        & This work                            \\
        \arrayrulecolor{black}\hline
      \end{tabular}%
    }%
    \endgroup
  }
\end{table}

\section{Related Work and Discussion}

Since Muon \citep{jordan2024muon} is a very efficient optimizer for functions of weight matrices, a lot of research has been put into, first, further improving its performance, and second, explaining its success.

\textbf{Improvements of Muon.} Regarding the first point, a large number of applications and improvements of Muon has been proposed in less than a year. \citet{liu2025muon} adapted the algorithm for training language models larger than NanoGPT. \citet{shah2025practical} organized efficient hyperparameter transfer by combining Muon with maximal update parametrization. To construct their COSMOS optimizer, \citet{chen2025cosmoshybridadaptive} applied computationally intensive updates of SOAP optimizer \citep{vyas2025soap} to a low-dimensional ``leading eigensubspace'' while using memory-efficient methods like Muon for the remaining parameters. \citet{amsel2025polar, grishina2025chebyshev} proposed more efficient alternatives to Newton-Schulz algorithm. \citet{si2025adamuon} introduced AdaMuon which combines element-wise adaptivity with orthogonal updates. We suppose that the described above or similar techniques can be applied to our optimizers as well. For example, F-Muon and S-Muon also benefit from faster alternatives to Newton-Schulz iterations, and Fanions may be a great substitute to Muon in COSMOS, because, as we have shown in \sectionref{sec:matrix-side}, Lanczos algorithm is much faster than Newton-Schulz iterations on very large matrices.

\textbf{Theory behind Muon.} Regarding the second point, there has been a prolonged gap in theory behind Muon, simplistic derivation of \citet{bernstein2025deriving} based on \citet{bernstein2024oldoptimizernewnorm} excluded. This gap, as it seems to us, is not even now completely closed. For example, \citet{kovalev2025understanding} has provided convergence guarantees of Muon and in various settings, from which, however, Muon's supremacy cannot be recovered. Indeed, although the obtained bounds depend on the norm choice, the asymptotics of the convergence remain the same as for NSGD and other optimizers, $K = \cO(\epsilon^{-4})$ in an $L$-smooth stochastic case.

A similar drawback has the article \citet{riabinin2025gluon}, where $L$-smoothness assumption is replaced with a more practical $(L_0, L_1)$-smoothness. By estimating smoothness and substituting it into their Theorem 1, the authors recovered the optimal fine-tuned stepsizes reported by \citet{pethick2025training}. However, they have not showed the optimality of the spectral or the RMS-to-RMS norm, which is observed in practice, as our comparison with NSGD highlights.

Common important drawback of the analyses is the consideration of the convergence by the norm of the gradient. As we showed in our experiments on CIFAR, the gradient norm may decrease only by a factor of ten, when the accuracy reaches 100\%.

We suppose that the reason for the recorded stark discrepancy between Neon and Muon performance, both of which are described by Stochastic Conditional Gradient \citep{pethick2025training} or Gluon frameworks, lies in the structure of the norm ball or in the preconditioner interpretation \citep{pethick2025trainingneuralnetworksscale}, which must be an object of further research.

\textbf{The LMO and Error Feedback.} We already mentioned that rank-$k$ unsharded Dion without error feedback and scaling of the update is Fanion-$k$. Since error feedback for Dion is very important as discovered by the ablation study in \citet{ahn2025dion}, F-Fanions and S-Fanions will benefit from it as well. In the context of federated learning, error feedback is effective even for compressed Muon \citep{gruntkowska2025efmuon}. The advantage of Fanions and S-Fanions is that less bits are required for transmission: $\sum_{i=1}^k \vu_i \vv_i^\top$ is easily transmitted as $\{\vu_1, \vv_1, \dots, \vu_k, \vv_k\}$ ($(m+n)\times k$ floats), while the sign part is as usual coded in $m \times n$ bits. Thus, the compression is ``built-in''. Moreover, there is an intriguing possibility to construct differentially private Fanions and S-Fanions with the specific more optimal non-Gaussian noise, as was done with DP-signSGD \citep{jang2024dplogsign}. This we leave for future research.

\textbf{The nuclear norm in the LMO.} As we found out only when transforming our results into an article, the nuclear norm has already been explored in the context of the linear minimization oracle. \citet{pethick2025sam} applied it to create $\nu$SAM, a new sharpness-aware minimization technique. It would be interesting to substitute $\normn{\cdot}$ with $\normkfk{\cdot}^\dagger$ in their approach. Since the SAM-neighborhood becomes more diverse, $k > 1$ might add to the accuracy boost, while preserving a small memory footprint and a small time overhead, if Dion-style power iterations are used.

\textbf{Orthogonal research.} Fanions, F-Fanions, and S-Fanions benefit from the general theoretical description of \citet{riabinin2025gluon}, and better learning rates can be predicted by calculating the trajectory smoothness. They could be transformed to Drop-Fanions by updating only some layers as in \citet{gruntkowska2025dropmuon}. They could be viewed as approximations of the Non-Euclidean Broximal Point Method for the corresponding norms \citep{gruntkowska2025noneuclideanbroximal}. One can clip them to produce ClippedScion-like algorithms \citep{pethick2025generalizedgradientnormclipping}. They could be made more memory-effective by the zero-order techniques that worked for Muon \citep{petrov2025zeromuon}. Finally, the results from \citet{shulgin2025inexactmuon} can be used to explain the robustness of Muon to the norm change that occurred in our experiments and to theoretically derive faster yet working approximate schemes to calculate the LMO; power iterations with a very limited number of iterations is a good algorithm to consider for the analysis.

\section{Conclusion}

In this article, we have generalized several successful algorithms, like Muon and Dion, to the LMO-based algorithms in the $\normkfk{\cdot}^\dagger$ norm. Also we have proposed the technique of ``regularizing'' the updates with NSGD, a trick to increase the robustness of the algorithms, which is motivated by the consideration of the $\normfstar{\cdot}$, $\normftwo{\cdot}$, and general $\norm{\cdot}_{\mathrm{F-KF-k}}$ norms. Generalizations of well-known norms and subsequent combinations of them may further improve performance of LMO-based algorithms. If a theory is developed that explains the discrepancy between performance of different algorithms based on the matrix norms, one will probably be able to intentionally construct norms optimal to given architectures and probably even their parameters.

We suggest that future works that deal with the non-Euclidean LMO explain in corollaries to their convergence theorems the empirical superiority of Muon to other Fanions. Without that, it is hard to believe that the bounds are relevant for practitioners.

\section{Author Contributions}

IO suggested using the nuclear norm in the \citet{bernstein2024oldoptimizernewnorm} framework. DM presented the problem at the MIPT optimization course and helped with editing the article. IK suggested using composite norms (though not the ones that induce F-Fanions and S-Fanions) and refactored \sectionref{sec:conic_combo}. NK suggested Lanczos algorithm as the most precise means to compute Fanions' updates, conducted experiments to prove this, and wrote \sectionref{sec:matrix-side}. AV conducted the finetuning of NanoGPT on Tiny Stories and wrote \sectionref{subsec:finetune}. All other work was done by AK: creating Fanions, F-Fanions, S-Fanions, experimenting on the Linear Least Squares, CIFAR-airbench and NanoGPT pretrain, and writing the manuscript.

@@

\appendix
\section{LMO for Neural Networks}

In a typical neural network, the objective function $F$ depends on a set of weight matrices $\{ \mW_1, \mW_2, \ldots \}$. The optimization framework we have described is applied in a layer-wise fashion. At each iteration $k$, a stochastic gradient $g(\mX^k, \xi^k)$ is computed using a mini-batch of data with the $\xi^k$ noise via backpropagation. This yields a separate gradient component, $\mG_i^k$, for each matrix $\mX_i$. The LMO-based update rule is then applied to each matrix $\mX_i$ using its corresponding gradient component $\mG_i^k$.

The update rule used in Muon optimizer is uSCG \citep{pethick2025training}. In the most general case, which involves momentum, it can be written as
\begin{equation}\label{eq:usgd_update}
  \begin{aligned}
    \mM^{k}     & = \alpha_{k} g(\mX^k, \xi^k) + (1 - \alpha_k)\mM^{k - 1}\,, \\
    \mX^{k + 1} & = \mX^k + \gamma_k\mathrm{LMO}(\mM^{k})\,.
  \end{aligned}
\end{equation}

\section{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}

Based on the lemma, we immediately find $\normfstar{\cdot}^\dagger$, which is related to F-Muon update. Indeed, after setting $\beta=1-\alpha$ and remembering that for smooth and bounded cases we can use $\min$ instead of $\inf$, we get
\begin{equation}\label{eq:normfstardual}
  \normfstar{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \norms{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
\end{equation}

If $\alpha = 1$, then $\mZ = \mY$, and we get $\normfstar{\mY}^\dagger = \norms{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normfstar{\mY}^\dagger = \normf{\mY}$.

Similarly, we find $\normftwo{\cdot}^\dagger$, which is related to F-Neon update:
\begin{equation}\label{eq:normftwodual}
  \normftwo{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \normn{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
\end{equation}

If $\alpha = 1$, then $\mZ = \mY$, and we get $\normftwo{\mY}^\dagger = \normn{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normftwo{\mY}^\dagger = \normf{\mY}$.

Unfortunately, $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$ do not have a closed-form expression.

\section{Visualization of different matrix norms}
\subsection{Duals to F* and F2 norms}

It follows from \lemmaref{lemma:dual_to_conv_comb} that the norm ball in $\normfstar{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\normn{\cdot}$ and $(1-\alpha)\normf{\cdot}$ and the norm ball in $\normftwo{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\norms{\cdot}$ and $(1-\alpha)\normf{\cdot}$.

In \figureref{fig:fduals} we plot these norms. On x-axis and y-axis, there are singular values $\sigma_1$, $\sigma_2$ respectively of a matrix from $\R^{m\times n}$ with $\min\{m,n\}=2$.

\begin{figure}[htbp]
\floatconts
  {fig:fduals}
  {\caption{Balls in the duals to F* and F2 norms for different $\alpha$}}
  {%
    \subfigure[lmo balls for F-Muon for different $\alpha$]{\label{fig:fstardual}%
      \includegraphics[width=0.48\linewidth]{figs/norms/fstardualball.pdf}}%
    \hfill
    \subfigure[lmo ball for F-Neon for different $\alpha$]{\label{fig:ftwodual}%
      \includegraphics[width=0.48\linewidth]{figs/norms/ftwodualball.pdf}}
  }
\end{figure}

\subsection{The Ky Fan norm and its dual}

1-balls in $l_\infty$, $l_1$ and $l_2$ norms are well-known from textbooks. But what about the Ky Fan $k$-norm? How can it be represented?

To showcase the complex structure of the Ky Fan $k$-norm and its dual, we suggest the illustrations \figureref{fig:kyfan_combined} with the ball in the Ky Fan $2$-norm in \figureref{fig:kyfan} and its dual in \figureref{fig:kyfandual}. On x-, y-, and z-axes, there are singular values $\sigma_1$, $\sigma_2$, and $\sigma_3$ respectively of a matrix from $\R^{m\times n}$ with $\min\{m,n\}=3$. In this particular case, we do not sort the singular values. In the proposed representation, we actually plot balls in the Top-$2$ norm $\max\{\abs{x}+\abs{y}, \abs{x}+\abs{z}, \abs{y}+\abs{z}\}$ and its dual norm $\max\{\max(\abs(x),\abs{y},\abs{z}), \frac{1}{2} (\abs{x}+\abs{y}+\abs{z})\}$. The resulting balls are much more complex than balls in $l_\infty$, $l_1$ and $l_2$ norms.

In fact, those balls can be described easier if we use the results from \citet{yu2012arithmetic}. The Ky Fan $2$-norm ball is an intersection of three $l_1$ balls in $(x,y)$, $(x, z)$, and $(y,z)$ spaces. The 1-ball in the dual Ky Fan $2$-norm is an intersection of 1-ball the in $l_\infty$ norm and $\frac{1}{2}$-ball in the $l_1$ norm.

\begin{figure}[htbp]
\floatconts
  {fig:kyfan_combined}
  {\caption{Ky Fan 2-norm and its dual}}
  {%
    \subfigure[The Ky Fan $2$-norm]{\label{fig:kyfan}%
      \includegraphics[width=0.45\linewidth]{figs/norms/KyFan.pdf}}%
    \hfill
    \subfigure[Dual to the Ky Fan $2$-norm]{\label{fig:kyfandual}%
      \includegraphics[width=0.45\linewidth]{figs/norms/KyFanDual.pdf}}
  }
\end{figure}

\section{More data for Linear Least Squares}
\label{sec:lls_plots_section}
